---
title: "Question 20 from Time Series Homework 3"
#output: html_notebook
---

Question 20:
In this problem, there is a way to do the problem using pencil-and-paper analysis, 
or you could even look up the answers in various time-series textbooks (paper or online). 
But, from my point of view, that would ruin the fun. 

i) Construct a simulation to experimentally determine the variance (or Std.Dev.) of an AR(1) process at time 10 with parameter a_1=0.4 and sigma=7. Use y(0)=0.

ii) Now try a wide variety of different sigma values. Make a record of your experiments. Describe the relationship between sigma or sigma^2 and Var(y(10)) or SD(y(10)), ideally using a formula.

iii) Now set sigma=7 and try a wide variety of a_1 values. Make a record of your experiments. Describe the relationship between a_1 and Var(y(10)) or SD(y(10)).

iv) Does y(100) or y(1000) behave differently than y(10) ?

v) -optional and hard- Now consider an AR(2) process. How does Var(y(1000)) vary with a_1 and a_2 ? Make a record of your experiments.


Try executing this chunk by clicking the *Run* button within the chunk or by placing your cursor inside it and pressing *Cmd+Shift+Enter*. 

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Cmd+Option+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Cmd+Shift+K* to preview the HTML file).

Start of Homework
```{r}
generate_AR1_process = function(y_1, a_1, sigma, size) {
  vector = 1:size
  vector[1] = 0
  for (i in 2:size) {
    vector[i] = a_1 * vector[i - 1] + sigma * rnorm(1, mean = 0, sd = 1)
  }
  return(vector)
}
```
The following block shows a line plot of the AR(1) process
```{r}
y_1 = 0
a_1 = 1.0
sigma = 7.0
size = 100
v = generate_AR1_process(y_1 = y_1, a_1 = a_1, sigma = sigma, size = size)
plot(v, type = "o", col = "red")
```
i) Construct a simulation to experimentally determine the variance (or Std.Dev.) of an AR(1) process at time 10 with parameter a_1=0.4 and sigma=7. Use y(0)=0.
```{r}
make_samples_of_y_n = function(y_1, a_1, sigma, size, iterations) {
  samples_of_y_n = sapply(1:ITERATIONS, 
                         function(x) generate_AR1_process(y_1 = y_1, a_1 = a_1, sigma = sigma, size = size)[size])
  return(samples_of_y_n)
}
y_1 = 0
a_1 = 1.0
sigma = 7.0
size = 10 # try with 100, 1000, etc
ITERATIONS = 100
samples_of_y_10 = make_samples_of_y_n(y_1 = y_1, a_1 = a_1, sigma = sigma, 
                                       size = size, iterations = ITERATIONS)
variance_of_samples = var(samples_of_y_10)
stdev_of_samples = sd(samples_of_y_10)
print(sprintf("variance of samples of y_%d: %5.3f", size, variance_of_samples))
print(sprintf("stdev of samples of y_%d: %5.3f", size, stdev_of_samples))
```

ii) Now try a wide variety of different sigma values. Make a record of your experiments. Describe the relationship between sigma or sigma^2 and Var(y(10)) or SD(y(10)), ideally using a formula.
```{r}
y_1 = 0
a_1 = 1.0
size = 10
sigmas = 1:20 # <- note that this is a range !! =)
ITERATIONS = 100
#calculates a range of varianes corresponding to the sigma range 1:10
the_variances_of_samples = sapply(sigmas, 
                                  function(sig) var(make_samples_of_y_n(y_1 = y_1,
                                                                         a_1 = a_1,
                                                                         sigma = sig,
                                                                         size = size,
                                                                         iterations = ITERATIONS)))
variance_output= "the variances (sigma: var): "
variance_output_result = sprintf("(sigma: %2d, var: %5.2f)\n", sigmas, the_variances_of_samples)
variance_output= cat(variance_output, "\n", variance_output_result)
plot(the_variances_of_samples, xlab="sigmas", type = "o", col = "red")

#calculates a range of standard deviations corresponding to the sigma range 1:10
the_stdev_of_samples = sapply(sigmas, 
                              function(sig) sd(make_samples_of_y_n(y_1 = y_1,
                                                                     a_1 = a_1,
                                                                     sigma = sig,
                                                                     size = size,
                                                                     iterations = ITERATIONS)))
stdev_output= "the stdevs (sigma: stdev): "
stdev_output_result = sprintf("(sigma: %2d, var: %5.2f)\n", sigmas, the_stdev_of_samples)
stdev_output= cat(stdev_output, "\n", stdev_output_result)
plot(the_stdev_of_samples, xlab="sigmas", type = "o", col = "red")
```

iii) Now set sigma=7 and try a wide variety of a_1 values. Make a record of your experiments. Describe the relationship between a_1 and Var(y(10)) or SD(y(10)).
```{r}
y_1 = 0
a_1s = seq(from = 0.1, to = 2.0, by = .1)
size = 10
sigmas = 7
ITERATIONS = 100

the_variances_of_samples = sapply(a_1s,
                                  function(a_1) var(make_samples_of_y_n(y_1 = y_1,
                                                                         a_1 = a_1,
                                                                         sigma = sigma,
                                                                         size = size,
                                                                         iterations = ITERATIONS)))
variance_output= "the variances (a_1: var): "
variance_output_result = sprintf("(a_1: %3.2f, var: %5.2f)\n", a_1s, the_variances_of_samples)
variance_output= cat(variance_output, "\n", variance_output_result)
plot(a_1s,the_variances_of_samples, xlab="a_1s", type = "o", col = "red")

#calculates a range of standard deviations corresponding to the sigma range 1:10
the_stdev_of_samples = sapply(a_1s,
                              function(a_1) sd(make_samples_of_y_n(y_1 = y_1,
                                                                     a_1 = a_1,
                                                                     sigma = sigma,
                                                                     size = size,
                                                                     iterations = ITERATIONS)))
stdev_output= "the stdevs (a_1: stdev): "
stdev_output_result = sprintf("(a_1: %3.2f, var: %5.2f)\n", a_1s, the_stdev_of_samples)
stdev_output= cat(stdev_output, "\n", stdev_output_result)
plot(a_1s, the_stdev_of_samples, xlab="a_1s", type = "o", col = "red")
```

Observation: as a_1 increases, the standard deviation and variance of the sample y_10s grow with exponential or polynomial growth (i.e. x^2).

--
iv) Does y(100) or y(1000) behave differently than y(10) ?

I calculated the variance and stdevs for y_10, y_100, and y_100 using the code block for answer (i)

for y_10
[1] "variance of samples of y_10: 470.086"
[1] "stdev of samples of y_10: 21.681"

for y_100
[1] "variance of samples of y_100: 5082.061"
[1] "stdev of samples of y_100: 71.289"

for y_1000
[1] "variance of samples of y_1000: 45840.570"
[1] "stdev of samples of y_1000: 214.104"

I tried plotting the variance and stdev for y_100, y_1000 with the sigmas from 1:20.
The standard deviations and variances increased in value but the trend in the increase was the same.

I also tried plotting the variance and stdevs for y_100, y_100 with a_1s ranging from .1 to 2.0.
The stdevs and variances are much greater for higher values of y.
The trend in growth seems to be the same for y_1 - y_10.  The growth looks exponential or polynomial (x^2)

I would say no, y_100 and y_1000 do not behave differently than y_10.

v) -optional and hard- Now consider an AR(2) process. How does Var(y(1000)) vary with a_1 and a_2 ? Make a record of your experiments.
